---
title: "rlesson3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(car)
library(corrplot)
library(PerformanceAnalytics)
library(questionr)
library(leaps)
#install.packages("PerformanceAnalytics")
```


```{r, results = F}
NHISdataAdult2018 <- read.csv("/Users/jaiveersingh/Desktop/samadult.csv")

head(NHISdataAdult2018)
names(NHISdataAdult2018)
str(NHISdataAdult2018)
```

```{r}
onlyAsianAndWhite <- NHISdataAdult2018[(NHISdataAdult2018$MRACBPI2 == 6 | NHISdataAdult2018$MRACBPI2 == 7 | NHISdataAdult2018$MRACBPI2 == 12 | NHISdataAdult2018$MRACBPI2 == 1), ] #Select only NHW, Asian Indians, Chinese, and Filipino

onlyAsianAndWhite$raceRecoded <- Recode(onlyAsianAndWhite$MRACBPI2, '"6" = "Chinese"; "7" = "Filipino"; "12" = "Asian Indian"; "1" = "NHW"')
```


```{r, results = F}
NHISdataAdult2018$DEP_1 # How often do you feel depressed
NHISdataAdult2018$DEP_2 # do you take dep medicine
NHISdataAdult2018$DEP_3R #how depressed were you the last time you felt depressed
NHISdataAdult2018$ALDURA17 #Duration of depresion
NHISdataAdult2018$AFLHCA17 #depression causes difficulty with activity
NHISdataAdult2018$ASISAD # so sad nothign cheers you up
NHISdataAdult2018$ASIHOPLS #hopeless
NHISdataAdult2018$ASIWTHLS #worthless
NHISdataAdult2018$ASINERV #felt nervous
NHISdataAdult2018$ASIRSTLS #felt restless
NHISdataAdult2018$ASIEFFRT #felt everything was an effort
NHISdataAdult2018$ASIMUCH #feelings interfered with life

NHISdataAdult2018$AHCAFYR2 # couldnt afford mental health counseling
NHISdataAdult2018$AHCSYR1 # have you seen mental health practicioner in past 12 months
```

linear regression

step 1: Identify Response and Predictors

response variable: ALDUDA17 - duration of depression

predictors: DEP_1, ASISAD, ASIHOPLS, ASIWTHLS, ASINERV, ASIRSTLS, ASIEFRT, ASIMUCH

step 2: check relationships with scatterplots
```{r}
duraCleaned <- onlyAsianAndWhite[onlyAsianAndWhite$ALDURA17 < 97, ]
duraCleaned$duradeplog <- log(duraCleaned$ALDURA17+1)

cp <- na.omit(duraCleaned[, c("duradeplog", "DEP_1", "ASISAD", "ASIHOPLS", "ASIWTHLS", "ASINERV", "ASIRSTLS", "ASIEFFRT", "ASIMUCH")])

sigcorr <- cor.mtest(cp, conf.level = .95)
corrplot.mixed(cor(cp), lower.col="black", upper = "ellipse", tl.col = "black", number.cex=.7, 
                tl.pos = "lt", tl.cex=.7, p.mat = sigcorr$p, sig.level = .05)
```

step 3: Fit Regression Model including all predictors
```{r}
lm1 <- lm(duradeplog ~ DEP_1 + ASISAD + ASIHOPLS + ASIWTHLS + ASINERV + ASIRSTLS + ASIEFFRT + ASIMUCH, data = cp)
summary(lm1)
```

Step 4: Identify significant predictors, remove insignificant predictors - backwards stepwise regression
Remove the predictor with the highest P values and rerun the linear model. Keep doing this and removing insignificant predictors until only significant predictors are left
```{r}
#Original - all predictors
lm2 <- lm(duradeplog ~ DEP_1 + ASISAD + ASIHOPLS + ASIWTHLS + ASIRSTLS + ASIEFFRT + ASIMUCH, data = cp)
summary(lm2)

#After all of the insignifcant predictors are removed
lm2 <- lm(duradeplog ~ ASIHOPLS + ASIRSTLS, data = cp)
summary(lm2)
```

Step 5: Check model assupmtions
```{r}
qqPlot(lm2, pch = 19)
```

Best Subsets Regression - Use model criteria to choose the best model (ex BIC or adjusted R squared)

```{r}
library(leaps)

#Get best subsets results - NOTE if we use . this means use all but the first column of the dataset
mod1 <- regsubsets(duradeplog ~ ., data = cp, nvmax = 7)

#What is in mod1?
names(mod1)

#use summary function to calculate comparison statistics
mod1sum <- summary(mod1)
names(mod1sum)

#notice that 'which' is actually a matrix with TRUE/FALSE information on inclusion of each possible predictor
mod1sum$which
```

```{r}
#Best model according to Adjusted R-squared
which.max(mod1sum$adjr2)

#Which variables are in model 5
names(cp)[mod1sum$which[which.max(mod1sum$adjr2), ]][-1]

#Fit this model and show results
cptemp <- cp[,mod1sum$which[which.max(mod1sum$adjr2), ]]
summary(lm(duradeplog ~ .,data = cptemp))

#Visual representation of results
plot(mod1, main = "Best Model According to Adjusted R-squared", scale="adjr2")
```

```{r}
#Best model according to Bayesian Information Criteria (BIC)
which.min(mod1sum$bic)

#Which variables are in model 2
names(cp)[mod1sum$which[which.min(mod1sum$bic), ]][-1]

#Fit this model and show results
cptemp <- cp[,mod1sum$which[which.min(mod1sum$bic), ]]
summary(lm(duradeplog ~ ., data = cptemp))

```

Logistic Regression
```{r}
seenPrac <- duraCleaned[!(is.na(duraCleaned$AHCSYR1)), ]
seenPrac <- duraCleaned[!(duraCleaned$AHCSYR1 == 7 | duraCleaned$AHCSYR1 == 8 | duraCleaned$AHCSYR1 == 9), ]
unique(seenPrac$AHCSYR1)
logm <- na.omit(seenPrac[, c("AHCSYR1", "duradeplog", "DEP_1", "ASISAD", "ASIHOPLS", "ASIWTHLS", "ASINERV", "ASIRSTLS", "ASIEFFRT", "ASIMUCH")])

lmod1 <- glm(as.factor(AHCSYR1) ~., data = logm, family = binomial("logit"))
summary(lmod1)
```

```{r}
x <- odds.ratio(lmod1)
row.names(x)
```

```{r}
boxLabels <- row.names(x)
yAxis = length(boxLabels):1
df <- data.frame(
  yAxis = length(boxLabels):1,
  boxOdds = x$OR,
  boxCILow = x$`2.5 %`,
  boxCIHigh = x$`97.5 %`
)

# Plot
p <- ggplot(df, aes(x = boxOdds, y = yAxis))
p + geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(aes(xmax = boxCIHigh, xmin = boxCILow), size = .5, height = .2, color = "gray50") +
  geom_point(size = 3.5, color = "orange") +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  scale_y_continuous(breaks = yAxis, labels = boxLabels) +
  scale_x_continuous(breaks = seq(0,7,1) ) +
  coord_trans(x = "log10") +
  ylab("") +
  xlab("Odds ratio (log scale)") +
  ggtitle("Have you seen a health practicioner in the last 12 months")
```











