---
title: "Introduction to Data Analysis in R (1)"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

#### Installing and Loading Packages

One of the great strengths of R is that it is open source. This means
that it is freely distributed and that its source code is open to
everyone, everywhere. This means that programmers from around the world
can contribute to the development of R by improving base R
functionalities or by adding new ones. A common way for R users to share
new functionalities that they have developed is through packages (also
called libraries) which are bundles of abstracted software (software
that is designed to be easy to read and use) that other users can
download and install.

To install packages in R, we use the `install.packages("package")`
command.

To load packages, we use `library(package)`.

Notice that when we load packages we don't need " " around the package
name. For this tutorial we will be using the
[tidyverse](https://www.tidyverse.org/) package, which is actually
multiple R packages bundled together.

```{r message = FALSE, warning = FALSE}
# install.packages("tidyverse")

library(tidyverse)
```

#### Loading Data

To read data into R, we can use functions from the `readr` package (part
of `tidyverse`). For example, you can read CSV
(**C**omma-**S**eparated-**V**alues) files using the `read_csv()`
function. Let's load the 2017 National Health Interview Survey (NHIS) as
we will use it in later sections.

```{r message = F, warning = F}
df.nhis <- read_csv("data/samadult_2017.csv")
```

## Common Variable Types

| Data Type          | Description                                         |
|--------------------|-----------------------------------------------------|
| `x <- TRUE`        | Logical = binary                                    |
| `x <- 15L`         | Integer = non-floating point numbers                |
| `x <- 15.0`        | Double = floating point with double precision       |
| `x <- "orange"`    | Character = just what it sounds like                |
| `x <- factor("a")` | Factor = a special type of coded (categorical) data |

You can switch between variable types using the functions as.\_\_\_ (for
example, `as.numeric()`, `as.factor()`). One exception is to use
`toString()` to convert to a String.

```{r}
# EX: SEX is encoded as a double, but if we wanted to convert it to a factor: 
df.nhis$SEX <- as.factor(df.nhis$SEX)
df.nhis$SEX

# Now with this factor we can learn more about its levels
```

Use `levels()` to see the distinct values of a factor.

```{r}
levels(df.nhis$SEX)
```

The order in which these levels appear tells us what number gets
assigned to each factor.

Use `as.numeric()` to see actual stored numeric values (the original
values in the dataframe, in our case).

```{r}
as.numeric(df.nhis$SEX)
```

To switch the order of the levels of the factor, use `relevel()`:

```{r}
# Set 2 as the reference level instead of 1
df.nhis$SEX <- relevel(df.nhis$SEX, "2")

# This will be more important when we talk about regression modelling!
```

## Data Types Overview

-   **Vectors** are one dimensional objects of length N, all one data
    type
-   **Lists** are an ordered lists of miscellaneous objects
-   **Data Frames** are rectangular arrays (rows and columns) of
    **mixed** data types (each column can be a different data type)

#### i. Vectors

all of same data type

```{r}
# To define a vector, you should just use vector <- c() and separate elements with a ,
a <- c("red", "blue")    # strings put into object b
b <- c(1,2,5,16,832)     # numbers put into object b
b[3]         # return the third element in this vector
b[c(1,3,5)]  # return the first, third, fifth elements
b[c(1:3)]    #return the first through third elements

# Exercise - for the vector below, rename the third element to "red" and print out the 4th element
test <- c("blue", "yellow", "green", "purple")
# Insert code here
```

#### ii. Lists

can be different data types

```{r}
x <- list("Red",45,c(1,2,5)) # put different variable types into list x
x[1]               #return first item from list
x[3]               #return third item from list as a list
x[[3]]             #return third item from list not as a list

class(x[3])               
class(x[[3]]) 

x.new <- xvec[[3]]     #put third item into a new object
```

#### iii. Tables

tables are a convenient way to organize data of a certain data type

```{r}
## Numerical Table - one way
table(df.nhis$SEX)

## Numerical Table - two way
table(df.nhis$MRACBPI2, df.nhis$SEX) #MRACBPI2 = race
```

```{r}
## Proportion Table! 
y <- table(df.nhis$MRACBPI2, df.nhis$SEX)

# The margin parameter represents whether you want to determine the proportions by row (1) or by column ()
prop.table(y, margin = 1) 
prop.table(y, margin = 2) 

# Exercise - create a proportion table between the variables MRACBPI2 and AGE_P, with proportions determined by column
# Insert code here
```

#### iv. Data Frames

This will be the data type that you use the most. It is the easiest way
to deal with large data in R. Each column can be its own data type. Most
functions require data frames as inputs.

```{r}
# df.nhis <- read_csv("samadult_2017.csv")
# Dataframes can be imported in R as done above 

# Or we can convert tables into dataframes
y.df <- as.data.frame(y)
```

To get the size/dimension/length of objects use

-   `dim()` for matrices and data frames
-   `length()` for vectors and lists

To find out the type of an object use

-   `str()` - this gives the structure of an object

```{r}
# See the first 5 rows of a dataframe
head(df.nhis.2)

# See the last 5 rows of a dataframe
tail(df.nhis.2)

# See all the columns of the dataframe
names(df.nhis.2)

# See the data types of all the variables
str(df.nhis.2)

#If you have a column that is numeric and want the 5 number summary
summary(df.nhis.2$AGE_P)
```

## 2. Working with Data Frames (Brief introduction to Tidyverse)

-   As a data scientist you will be spending 70 \~ 90 % of your time
    pre-processing data (i.e. cleaning, re-coding, re-formatting).
-   Using indexing and other traditional ways of formatting data works
    BUT the code can get really long and complex, which increases the
    chances of error.
-   R is a old language (28 years old!), so most of the new innovation
    occurs in the shape of software packages. There is a coherent system
    of packages for data manipulation, exploration, and visualization
    called `tidyverse` that will make your life easier.
-   We will be covering `dplyr` and `ggplot2` (both are part of the
    `tidyverse` ecosystem) in this tutorial.

The `dplyr` package aims to provide a function for each basic verb of
data manipulation. I will introduce the some of the most heavily used
"verbs" (i.e. functions) but there are many more available. These are
some resources you can use to learn more about `dplr`.

-   [dplyr cheet
    sheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

-   [R for data science](https://r4ds.had.co.nz/)

#### i. Renaming Columns

For this tutorial we will be using the 2017 wave of the National Health
Interview Survey (NHIS). One of the first things you'll notice about
NHIS is that columns have really obscure names. For example, the column
that contains the race/ethnicity info is called **MRACBPI2** (Don't ask
me why). So, one the first things you'll want to do is change the column
name.

```{r}
### Make a copy of the NHIS data frame ###
df.nhis.copy <- df.nhis

### Traditional way of doing it (not recommended) ###

nhis.colnames <- names(df.nhis.copy) # A vector of column names in the dataframe
which(nhis.colnames == "MRACBPI2") # Figure out the column number of MRACBPI2
nhis.colnames[12] <- "race" # Change "MRACBPI2" to "race"
names(df.nhis.copy) <- nhis.colnames # Rename the column in the dataframe

df.nhis.copy$race # Check

# names(df.nhis.copy)[names(df.nhis.copy) == "MRACBPI2"] <- "race" # A one liner

### The dplyr way of doing it ###
df.nhis <- dplyr::rename(df.nhis, race = MRACBPI2) # There may be other functions named rename
                                                   # thus I'm specifying that I want dplyr's
                                                   # rename function using dplyr::
df.nhis$race # Check

# I can also rename multiple columns
df.nhis <- dplyr::rename(df.nhis,
                         age = AGE_P,
                         sex = SEX,
                         smoke = SMKSTAT2,
                         alcohol = ALCSTAT)

# I can also do this (More advanced but may come in handy)
old.names <- c("STREV", "CANEV", "EPHEV", "DIBEV1", "HYPEV")           # The index of names
new.names <- c("stroke", "cancer", "lung", "diabetes", "hypertension") # must match

df.nhis <- rename_at(df.nhis, vars(old.names), ~new.names)

# To see the manual for "rename" you can type
?rename
```

#### ii. Selecting Columns

Something else you may have noticed is that the NHIS data set contains
702 columns. That amount of columns is way too much for any human being
to handle and most of them are probably not going to be related to your
research question for the summer. We can select a subset of columns that
we are interested in using `select()`.

```{r}
### Traditional way ###
# select a single column
df.nhis[,"age"]

# select multiple columns
df.nhis[,c("age", "sex", "race", "stroke", "cancer")]

### dyplr way ###
# select a single column
select(df.nhis, age)

# select multiple columns
select(df.nhis, age, sex)

# EXERCISE: Select the columns for age, sex, race, smoke, stroke, diabetes, and hypertension
#           and save it in a dataframe called df.nhis.2 

df.nhis.2 <- 
  
# Again, you can read the documentation for "select" by doing:
?select
```

#### iii. Creating New Columns and Re-coding Variables

Let's take a look at the sex column for a second.

```{r}
df.nhis.2$sex # Extract a column as a vector
```

It seems like the sex variable is a bunch of 1s and 2s. From the NHIS
code book we know that 1 represents male and 2 represents female. We
want to re-code this variable so that we have "males" and "females"
instead of 1s and 2s. We can do this by creating a new column using
`mutate()` in combination with `ifelse`. I will also introduce a really
powerful tool: `%>%` (pipe function).

```{r}
(
df.nhis.2 <- df.nhis.2 %>% 
  mutate(
    r_sex = ifelse(sex == 1, "male", "female"),
  )
)
```

Here, we use `mutate()` to create a new column named `r_sex` (short for
re-coded sex) where if the value of the `sex` column has a value of "1",
then `r_sex` would have the "male". Else, `r_sex` would be coded as
"female".

The pipe function passes the data to the next function, sort of like how
a ball would get passed around to different players in soccer or basket
ball. Here, `%>%` passes the contents of `df.nhis.2` the mutate
function. Note that we no longer have to include `df.nhis.2` in the
first argument of `mutate()`. The pipe is a powerful tool because it
allows you to focus on the transformations, not what's being
transformed, which make the code easier to read. It will help you chain
a series data transformation operations to create a data wrangling
"sentence".

```{r}
# EXERCISE: Create a new column named r_diabetes, that recodes the diabetes column
#           Hint: 1 -> "yes", 2 -> "no", 3 ~> Ignore for now!

df.nhis.2 <- df.nhis.2 %>% 
  mutate(
    r_diabetes = ifelse(diabetes == 1, "yes", "no")
  )
```

Now that we have re-coded `sex` and `diabetes` let's try and re-code
`race`. The `ifelse` verb worked well on the previous two variables
because they were binary, but `race` has multiple categories (well
duh...) so we need something else. Fortunately, there is a generalized
if-else verb called `case_when()`.

```{r}
# Using mutate and case_when to recode race
df.nhis.2 <- df.nhis.2 %>%
  mutate(
    r_race = case_when(
      race == 1 ~ "nhw",
      race == 6 ~ "chinese",
      race == 7 ~ "filipino",
      race == 12 ~ "asian indian",
      TRUE ~ "other"  # Handles all other cases!
    ),
    r_race = factor(r_race, levels = c("nhw", "chinese", "filipino", "asian indian"))
  )

# While we're at it let's also stratify the age variable
df.nhis.2 <- df.nhis.2 %>%
  mutate(
    r_age = case_when(
      age < 26 ~ "18~25",
      age < 45 ~ "26~44",
      age < 65 ~ "45~64",
      age >= 65 ~ "+65"
    ),
    r_age = factor(r_age, levels = c("18~25","26~44","45~64","+65"))
  )

# A fancy way of doing it for advanced users
df.nhis.2 %>%
  mutate(
    r_race = plyr::mapvalues(race,
                             from = c(1,6,7,12),
                             to = c("nhw", "chinese", "filipino", "asian indian"))
  )

# EXERCISE: Using mutate() and case_when() recode the diabetes column again, but this time:
#           1 -> "yes", 2 -> "no", 3 -> "prediabetes"
df.nhis.2 <- df.nhis.2 %>%
  mutate(
    r_diab = case_when(
      diabetes == 1 ~ "yes",
      diabetes == 2 ~ "no",
      diabetes == 3 ~ "prediabetes"
    )
  )
```

#### iv. Filtering rows

Sometimes (a lot of times really), you want to limit your sample to
respondents with a certain characteristic. For example, this program
focuses on Asian American subgroups so we might want to filter out the
other race groups from our sample (Not that they are not important!).
For this, we can use the `filter()` verb.

```{r}
# Filter the data for nhw, and AA subgroups
df.nhis.2 %>% 
  filter(r_race %in% c("nhw", "chinese", "filipino", "asian indian")) # %in% = if included in

# We can also do
df.nhis.2 %>%
  filter(r_race != "other") # != means "not equal to"

# EXERCISE: Filter out those responents who have prediabetes
# df.nhis.2 %>% filter(...Your code here...)
df.nhis.2 <- df.nhis.2 %>% 
  filter(r_diab != "prediabetes")
```

#### v. Creating Summaries by Group

In the previous section, we learned about how to create bi-variate
summary tables using the `table()` and `prop.table()` functions. For
example we created a table that had the counts for number of respondents
in each race/ethnicity group within each age group. If we were to do
this by hand, we would probably take the following steps:

1.  Splitting the dataset into groups (e.g. age, race/ethnicity)
    (**Split**)
2.  Count the number of respondents in each group (**Apply**)
3.  Combine the results into a summary table (**Combine**)

Many data analysis tasks can be approached using this
split-apply-combine paradigm, and `dplyr` makes this very easy with
`group_by()` and `summarise()` functions. Let's begin with a simple
example where we count the number of respondents in each race/ethnicity
group.

```{r}
df.nhis.2 %>% 
  group_by(r_race) %>%
  summarise(count = n())
```

Here, `group_by()` changed the unit of analysis from the complete data
set to individual race/ethnicity groups. The `summarise()` function then
collapses each data group into single rows to create a table of total
counts. The `n()` function summarizes the data by computing the counts,
and therefore it belongs to a family of functions called **summary
functions**. Some useful summary functions include:

-   `mean()`
-   `median()`,
-   `sum()`
-   `min()`,
-   `max()`,
-   `sd()`: (standard deviation)

```{r}
# EXERCISE: Create a table with the median age for each race/ethnicity
df.nhis.2 %>%
  group_by(r_race) %>%
  summarise(avg = median(age, na.rm = T))
```

We can also group by multiple variables (e.g. race/ethnicity & age
group). When you do this each summary peels off one level of the
grouping, making it easy to progressively roll up a data set.

```{r}
df.nhis.2 %>% 
  group_by(r_race, r_age) %>% 
  summarise(count = n())

# EXERCISE: Using group_by() and summarise() compute the number of people with diabetes in each race group
df.nhis.2 %>%
  group_by(r_race, r_diab) %>%
  summarise(count = n())
```

**Ungrouping**

If you need to remove grouping, and and return to operations on the
ungrouped data, use `ungroup()`.

```{r}
df.nhis.2 %>%
  group_by(r_race, r_age) %>%
  summarise(count = n()) %>%
  ungroup() # This removes the grouping
```

**Grouped mutates**

Grouping is most useful in conjunction with `summarise()`, but we can
also do convenient operations with `mutate()`

In the previous exercise, we computed the number of diabetic and
non-diabetic respondents in each race/ethnicity group. The natural next
step would be to obtain the proportions (i.e. the percentage of diabetic
and non-diabetic respondents for each race).

```{r}
( # Repeat EXERCISE
df.diab <- df.nhis.2 %>%
  group_by(r_race, r_diab) %>% 
  summarise(count = n()) %>%
  ungroup()
)

df.diab %>%
  group_by(r_race) %>%
  mutate(
    race_count = sum(count),
    percentage = count/race_count * 100
  )
```

By applying a `group_by()` before `mutate()`, the summary function
`sum()` will compute the total counts for each race/ethnicity group
instead of the complete data. We can then use this to compute the
percentage of people with diabetes for each race group.

## 3. Introduction to Data Visualization with ggplot2

For your in-class projects this week you will be exploring the data sets
in order to gain a better understanding of them and to start developing
your research question. You will be presenting these preliminary results
to the implementation fellows, TAs, as well as Drs. Srinivasan, and
Palaniappan. To maximally capitalize our time together, you should make
sure that your data is presented with graphics that are easy to read and
understand. This section will cover the basics of\`ggplot2\` and how to
create some basic plots that you will be using heavily this week.

`ggplot2` is a R package for producing statistical graphics and is
written by [Hadley Wickham](http://hadley.nz), a giant for modern
statistical computing in R. Unlike other graphics packages, `ggplot2`
has an underlying
[grammar](https://www.springer.com/gp/book/9780387245447) which makes it
easy to learn and intuitive to work with. I will only be going over the
very basics here, but if you want to master \`ggplot2\` Dr. Wickham has
released his [book](https://ggplot2-book.org/introduction.html) for free
online.

#### i. First Steps

![The 7 layers of
ggplot2](https://raw.githubusercontent.com/ShozenD/care-tutorials-2021/main/figures/ggplot-layered.png){width="367"}

When you create plots using \`ggplot2\` you want to think in terms of
"layers". The figure above gives you a very good image of each layer and
their ordering. Let's start with the three most important layers at the
bottom.

Every `ggplot2` plot has three key components:

1.  **data**. The foundation of your plot and always the first argument
    of the `ggplot` function.
2.  A set of **aesthetic mappings** between variables in the data and
    visual properties (e.g. x-axis, y-axis, color, shape, size).
    Aesthetic mappings are defined using the `aes` function.
3.  At least one geometric layer which describes how to render (i.e.
    draw) each observation. Layers are usually created with a `geom`
    function.

```{r}
head(iris)

ggplot(iris, aes(x = Sepal.Length, y = Petal.Length)) + 
  geom_point()

# EXERCISE: Create a scatter plot of Sepal.Length and Sepal.Width and color it by Species
```

Pay attention to the structure of program (i.e. function call): data and
aesthetic mappings are supplied to `ggplot()`, then geometric layers are
added with `+`. This is the basic pattern for every `ggplot2` program.

```{r}
# EXERCISE: Lets create some plots using df.nhis.2

# 1. A histogram of age. hint: Use geom_histogram()
ggplot(df.nhis.2, aes(x = age)) + 
  geom_histogram()

# 2. Boxplots with recoded race (r_race) on the x-axis and age on the y-axis. 
# hint: Use geom_boxplot()
ggplot(df.nhis.2, aes(x = r_race, y = age)) + 
  geom_boxplot()

# 3. A barplot for race counts (r_race). hint: geom_bar()
ggplot(df.nhis.2, aes(r_race)) + 
  geom_bar()
```

#### ii. Stacked Bar Plots

Now that we have the basics down, we can begin to use `ggplot` to
visualize our data. In a previous section you learned about how to make
bi-variate summary tables. For example, if we want to find the number of
respondents with diabetes within each race/ethnicity group we can easily
create a table by:

```{r}
table(df.nhis.2$r_race, df.nhis.2$r_diab)
```

However, when you present this data you want use plots not tables. To do
this we need to go through the following two steps.

1.  Count the number of obs. in each group.
2.  Create a stacked bar plot

Recall that we can count obs. in groups using the `group_by()` and
`summarise()` functions.

```{r}
df.diab <- df.nhis.2 %>% 
  group_by(r_race, r_diab) %>% 
  summarise(n = n())
```

We can then pass this data to ggplot2 for plotting. We want
race/ethnicity to be on the x-axis, the number of counts on the y-axis,
and we want to color the bars by diabetes status. Recall that to create
a bar plot we use the `geom_bar()` function. In order to create a
stacked bar plot, within the `geom_bar()` function we need to set
`position = "fill"` and `stat = "identity"` as such:

```{r}
ggplot(df.diab, aes(x = r_race, y = n, fill = r_diab)) + 
  geom_bar(
    position = "fill", 
    stat = "identity"
  )
```

Now, instead of absolute counts we can see the proportion of people with
diabetes in each race/ethnicity group! But we can't show this plot to
anyone yet because it doesn't have a title, the y-axis has the wrong
label, and the legend has a terrible name. We can customize the labels
using the `labs()` function.

```{r}
ggplot(df.diab, aes(x = r_race, y = n, fill = r_diab)) + 
  geom_bar(
    position = "fill", 
    stat = "identity"
  ) + 
  labs(
    x = "Race/Ethnicity",
    y = "Proportion",
    fill = "Diabetes Status"
  )
```

If you still feel like the plot is not to your liking you can make it
fancier by adding other options to `geom_bar()`, and by adding other
layers to the plot. We don't have time to cover everything today but
here is a example:

```{r}
ggplot(df.diab, aes(x = r_race, y = n, fill = r_diab)) + 
  geom_bar(
    width = 0.5,
    position = "fill", 
    stat = "identity"
  ) + 
  labs(
    x = "Race/Ethnicity",
    y = "Proportion",
    fill = "Diabetic"
  ) +
  scale_fill_manual(values = c("#488f31", "#de425b")) + 
  theme_minimal() + 
  theme(aspect.ratio = 1)
```

## I'm Bored! Gimme something to do!

#### Level 1: Adding Annotation

The main issue with stacked bar plots is that we are dropping sample
size information. Figure out a way to annotate stack bar plots with the
sample size of each group (e.g. race/ethnicity).

**Hint**:

-   Create another data frame with the sample sizes for each group.
-   You can use `geom_text()` to annotate plots.

```{r}

```

#### Level 2: A Generalized Function for Stacked Bar Plots

Since you'll probably be making lots and lots of stacked bar plots this
week, it will be extremely useful to define a function that takes any
pair of variable (given that they are both categorical) from NHIS and
generates a nice looking plot. Complete the following R function.

**Useful Resources**:

-   **Chapter 3 and 4** of [Tidy
    evaluation](https://tidyeval.tidyverse.org/) by Lionel Henry and
    Hadley Wickham

```{r}
#' Creates a stacked bar plot for a given pair of categorical NHIS variables.
#'
#' @param data A NHIS data frame
#' @param x Variable name
#' @param y Variable name
#'
#' @return A stacked bar plot created using ggplot2
#' @export
#'
#' @examples
#' \dontrun{
#'  ggstackbar(df.nhis, race, diabetes)
#' }
ggstackbar <- function(data, x, y){
  # Data wrangling
  
  
  # Plot the data
  
}
```

## Homework

**Kessler 6 Distress Scale**

The NHIS measures general psychological distress using a well-validated
[Likert-type scale](https://en.wikipedia.org/wiki/Likert_scale) called
the Kessler 6 psychological distress scale (abbr. K-6). Below is a list
of the 6 questions within the K-6:

1.  During the past 30 days, how often did you feel so sad that nothing
    could cheer you up?
2.  During the past 30 days, how often did you feel nervous?
3.  During the past 30 days, how often did you feel restless or fidgety?
4.  During the past 30 days, how often did you feel hopeless?
5.  During the past 30 days, how often did you feel that everything was
    an effort
6.  During the past 30 days, how often did you feel worthless?

#### Assignment 1

Load the 2017 wave of the NHIS sample adult file and using the code
book, identify the variable names corresponding to K-6. **Rename** the
corresponding column names to `sad`, `nervous`, `restless`, `hopeless`,
`effort`, and `worthless` and create a subset of the data by
**selecting** only the 6 columns and saving the resulting dataframe to a
variable named `df.k6`.

Hint: Use `rename()` and `select()` from `dplyr`.

```{r}

```

#### Assignment 2

There is an issue with the NHIS coding of the K-6 answers. Currently,
the answer "ALL of the time" is encoded as 1 while, "NONE of the time"
is encoded as 5 (refer to code book). However, for the purpose of our
analysis, we want "ALL of the time" to be encoded as 4, "MOST of the
time" to be encoded as 3, ..., and "NONE of the time" to be encoded as
0. Additionally, we want to exclude those who "Refused", "Not
ascertained", or "Don't know". Write the code that such that each K-6
variable is scored on a range of 0-4 (higher score indicating more
distress) and filter out those respondents who did not or could not
answer.

Hint: Use `mutate()` in combination with `case_when()` or `ifelse()`. Use `filter()` to exclude those with certain responses.

```{r}

```

#### Assignment 3

The overall K-6 scale ranges for 0-24, with higher score indicating more severe levels of distress. Generally, the K-6 is stratified into 4 levels with 0 indicating no distress, a score of less than 6 indicating mild distress, a score of less than 12 indicating moderate distress, and anything greater or equal to 12 indicating severe distress. Write the code that adds the K-6 scores to generate a total score and then stratify this into 4 categories.

Hint: You can add columns in `mutate()` (e.g. `mutate(total = colA + colB + colC)`). Use mutate in combination with `case_when()` to stratify.

```{r}

```

#### Assignment 4

Use a stacked bar chart to visualize the proportion of people for each race/ethnic subgroup (NHW, Chinese, Asian Indian, and Filipino) who experienced none, mild, moderate, or severe psychological distress.

```{r}

```

