---
title: "Introduction to Data Analysis in R (1)"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing and Loading Packages

```{r message = FALSE, warning = FALSE}
# install.packages("tidyverse")
# install.packages("car")

library(tidyverse)
```

```{r message = F, warning = F}
df.nhis <- read_csv("data/samadult_2017.csv")
```

## Common Variable Types
Data Type | Description
----------- | ---------------------------------------------------------------------------
`x <- TRUE`  | Logical = binary
`x <- 15L`   | Integer = non-floating point numbers
`x <- 15.0` | Double = floating point with double precision
`x <- "orange"` | Character = just what it sounds like
`x <- factor("a")` | Factor = a special type of coded (categorical) data

You can switch between variable types using the functions as.___ (for example, `as.numeric()`, `as.factor()`). One exception is to use `toString()` to convert to a String. 

```{r}
# EX: SEX is encoded as a double, but if we wanted to convert it to a factor: 
df.nhis$SEX <- as.factor(df.nhis$SEX)
df.nhis$SEX

# Now with this factor we can learn more about its levels
```

Use `levels()` to see the distinct values of a factor.

```{r}
levels(df.nhis$SEX)
```

The order in which these levels appear tells us what number gets assigned to each factor.

Use `as.numeric()` to see actual stored numeric values (the original values in the dataframe, in our case).
```{r}
as.numeric(df.nhis$SEX)
```

To switch the order of the levels of the factor, use `relevel()`: 
```{r}
# Set 2 as the reference level instead of 1
df.nhis$SEX <- relevel(df.nhis$SEX, "2")

# This will be more important when we talk about regression modelling!
```

## Data Types Overview

-   **Vectors** are one dimensional objects of length N, all one data
    type
-   **Lists** are an ordered lists of miscellaneous objects
-   **Data Frames** are rectangular arrays (rows and columns) of
    **mixed** data types (each column can be a different data type)

#### Vectors

all of same data type

```{r}
# To define a vector, you should just use vector <- c() and separate elements with a ,
a <- c("red", "blue")    # strings put into object b
b <- c(1,2,5,16,832)     # numbers put into object b
b[3]         # return the third element in this vector
b[c(1,3,5)]  # return the first, third, fifth elements
b[c(1:3)]    #return the first through third elements

# Exercise - for the vector below, rename the third element to "red" and print out the 4th element
test <- c("blue", "yellow", "green", "purple")
# Insert code here
```

#### Lists

can be different data types

```{r}
x <- list("Red",45,c(1,2,5)) # put different variable types into list x
x[1]               #return first item from list
x[3]               #return third item from list as a list
x[[3]]             #return third item from list not as a list

class(x[3])               
class(x[[3]]) 

x.new <- xvec[[3]]     #put third item into a new object
```

#### Tables

tables are a convenient way to organize data of a certain data type

```{r}
## Numerical Table - one way
table(df.nhis$SEX)

## Numerical Table - two way
table(df.nhis$MRACBPI2, df.nhis$SEX) #MRACBPI2 = race
```

```{r}
## Proportion Table! 
y <- table(df.nhis$MRACBPI2, df.nhis$SEX)

# The margin parameter represents whether you want to determine the proportions by row (1) or by column ()
prop.table(y, margin = 1) 
prop.table(y, margin = 2) 

# Exercise - create a proportion table between the variables MRACBPI2 and AGE_P, with proportions determined by column
# Insert code here
```

#### Dataframes!

This will be the datatype that you use the most. It is the easiest way
to deal with large data in R. Each column can be its own data type. Most
functions require dataframes as inputs.

```{r}
# df.nhis <- read_csv("samadult_2017.csv")
# Dataframes can be imported in R as done above 

# Or we can convert tables into dataframes
y.df <- as.data.frame(y)
```

To get the size/dimension/length of objects use

-   `dim()` for matrices and dataframes
-   `length()` for vectors and lists

To find out the type of an object use

-   `str()` - this gives the structure of an object

```{r}
# See the first 5 rows of a dataframe
head(df.nhis.2)

# See the last 5 rows of a dataframe
tail(df.nhis.2)

# See all the columns of the dataframe
names(df.nhis.2)

# See the data types of all the variables
str(df.nhis.2)

#If you have a column that is numeric and want the 5 number summary
summary(df.nhis.2$AGE_P)
```

## Working with Data Frames (Brief introduction to Tidyverse)

-   As a data scientist you will be spending 70 \~ 90 % of the time
    pre-processing data (i.e. cleaning, re-coding, re-formatting).
-   Using indexing and other traditional ways of formatting data works
    BUT the code can get really long and complex, which increases the
    chances of error.
-   R is a old language (28 years old!), so most of the new innovation
    occurs in the shape of software packages. There is a coherent system
    of packages for data manipulation, exploration, and visualization
    called `tidyverse` that will make your life much easier.
-   We will be covering `dplyr` and `ggplot2` (both are part of the
    `tidyverse` ecosystem) in this tutorial.

`dplyr` aims to provide a function for each basic verb of data
manipulation. I will introduce the 3 most heavily used "verbs" but if
you want to know what else is available, check out this [cheet
sheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).

#### Renaming Columns

One of the first things you'll notice about NHIS is that some columns
have really obscure column names. For example, the column that contains
the race/ethnicity info is called **MRACBPI2** (Don't ask me why). So,
one the first things you'll probably want to do is change the name!

```{r}
### Make a copy of the NHIS data frame ###
df.nhis.copy <- df.nhis

### Traditional way of doing it (not recommended) ###

nhis.colnames <- names(df.nhis.copy) # A vector of column names in the dataframe
which(nhis.colnames == "MRACBPI2") # Figure out the column number of MRACBPI2
nhis.colnames[12] <- "race" # Change "MRACBPI2" to "race"
names(df.nhis.copy) <- nhis.colnames # Rename the column in the dataframe

df.nhis.copy$race # Check

# names(df.nhis.copy)[names(df.nhis.copy) == "MRACBPI2"] <- "race" # A one liner

### The dplyr way of doing it ###
df.nhis <- dplyr::rename(df.nhis, race = MRACBPI2) # There may be other functions named rename
                                                   # thus I'm specifying that I want dplyr's
                                                   # rename function using dplyr::
df.nhis$race # Check

# I can also rename multiple columns
df.nhis <- dplyr::rename(df.nhis,
                         age = AGE_P,
                         sex = SEX,
                         smoke = SMKSTAT2,
                         alcohol = ALCSTAT)

# I can also do this (More advanced but may come in handy)
old.names <- c("STREV", "CANEV", "EPHEV", "DIBEV1", "HYPEV")           # The index of names
new.names <- c("stroke", "cancer", "lung", "diabetes", "hypertension") # must match

df.nhis <- rename_at(df.nhis, vars(old.names), ~new.names)

# To see the manual for "rename" you can type
?rename
```

#### Selecting Columns

Another thing you may be thinking now is that 702 columns is too much! I
don't need all the variables, I only need the ones that I am interested
in! To select columns, simply use the `select` verb (function).

```{r}
### Traditional way ###
# select a single column
df.nhis[,"age"]

# select multiple columns
df.nhis[,c("age", "sex", "race", "stroke", "cancer")]

### dyplr way ###
# select a single column
select(df.nhis, age)

# select multiple columns
select(df.nhis, age, sex)

# EXERCISE: Select the columns for age, sex, race, smoke, stroke, diabetes, and hypertension
#           and save it in a dataframe called df.nhis.2 
df.nhis.2 <- df.nhis %>% 
  select(age, sex, race, smoke, stroke, diabetes, hypertension)

# df.nhis.2 <- (Uncomment and enter your code here)
  
# Again, you can read the documentation for "select" by doing:
?select
```

#### Creating New Columns and Re-coding Variables

Let's take a look at the sex column for a second.

```{r}
df.nhis.2$sex # Extract a column as a vector
```

It seems like the sex variable is a bunch of 1s and 2s. From the code
book we know that 1 means male and 2 means female, but this is anyone
without a code book won't know this and it is easy to forget. We want to
re-code this variable such that instead of 1s and 2s, we have "males"
and "females". We can do this by creating a new column using the
`mutate` verb in combination with `ifelse`. I will also introduce a
really convenient function `%>%` called a pipe function.

```{r}
df.nhis.2 %>% 
  mutate(
    r_sex = ifelse(sex == 1, "male", "female"),
  )

# Without %>%
df.nhis.2 <- mutate(df.nhis.2, 
                    r_sex = ifelse(sex == 1, "male", "female"))
```

The pipe function passes the data frame to the mutate function, which
means we don't have to type `mutate(df.nhis.2, ...)`. I will demonstrate
later how using the pipe function will make your code much more shorter
and readable.

Here, we use `mutate` to create a new column named `r_sex` (short for
re-coded sex) where if the value of the `sex` column was "1", then
`r_sex` would be coded as "male". Else, `r_sex` would be coded as
"female".

```{r}
# EXERCISE: Create a new column named r_diabetes, that recodes the diabetes column
#           Hint: 1 -> "yes", 2 -> "no", 3 ~> Ignore for now!

# df.nhis.2 <- (Uncomment and enter your code here!)
```

Now that we have re-coded `sex` and `diabetes` let's try and re-code
`race`. The `ifelse` verb worked well on the previous two variables
because they are binary, but `race` has multiple categories (well
duh...) so we need something else. Fortunately, there is a generalized
if-else verb called `case_when`.

```{r}
# Using mutate and case_when to recode race
df.nhis.2 <- df.nhis.2 %>%
  mutate(
    r_race = case_when(
      race == 1 ~ "nhw",
      race == 6 ~ "chinese",
      race == 7 ~ "filipino",
      race == 12 ~ "asian indian",
      TRUE ~ "other"  # Handles all other cases!
    ),
    r_race = factor(r_race, levels = c("nhw", "chinese", "filipino", "asian indian"))
  )

# While we're at it let's also stratify the age variable
df.nhis.2 <- df.nhis.2 %>%
  mutate(
    r_age = case_when(
      age < 26 ~ "18~25",
      age < 45 ~ "26~44",
      age < 65 ~ "45~64",
      age >= 65 ~ "+65"
    ),
    r_age = factor(r_age, levels = c("18~25","26~44","45~64","+65"))
  )

# A fancy way of doing it for advanced users
df.nhis.2 %>%
  mutate(
    r_race = plyr::mapvalues(race,
                             from = c(1,6,7,12),
                             to = c("nhw", "chinese", "filipino", "asian indian"))
  )

# EXERCISE: Using mutate() and case_when() recode the diabetes column again, but this time:
#           1 -> "yes", 2 -> "no", 3 -> "prediabetes"
df.nhis.2 <- df.nhis.2 %>%
  mutate(
    r_diab = case_when(
      diabetes == 1 ~ "yes",
      diabetes == 2 ~ "no",
      diabetes == 3 ~ "prediabetes"
    )
  )
```

#### Filtering rows

Sometimes (a lot of times really), you want to limit your sample to
respondents with a certain characteristic. For example, this program
focuses on Asian American subgroups so we might want to filter out the
other race groups from our sample (Not that they are not important!).
For this, we can use the `filter` verb.

```{r}
# Filter the data for nhw, and AA subgroups
df.nhis.2 %>% 
  filter(r_race %in% c("nhw", "chinese", "filipino", "asian indian")) # %in% = if included in

# Use can also do
df.nhis.2 %>%
  filter(r_race != "other") # != means "not equal to"

# EXERCISE: Filter out those responents who have prediabetes
# df.nhis.2 %>% filter(...Your code here...)
df.nhis.2 <- df.nhis.2 %>% 
  filter(r_diab != "prediabetes")
```

#### Creating Summaries by Group

In the previous section, we learned about how to create bi-variate
summary tables using the `table()` and `prop.table()` functions. For
example we created a table that had the counts for each race/ethnicity
for each age group. If we were to do this by hand, we would probably
begin by:

1.  Splitting the dataset into groups (age, race/ethnicity) (**Split**)
2.  Count the number of respondents in each group (**Apply**)
3.  Combine the results into a summary table (**Combine**)

Many data analysis tasks can be approached using this
split-apply-combine paradigm, and `dplyr` makes this every easy through
the use of `group_by()` and `summarise()` functions. Let's begin with an
easy example where we count the number of repondents in each
race/ethnicity group.

```{r}
df.nhis.2 %>% 
  group_by(r_race) %>%
  summarise(count = n())
```

Here, we used `group_by(r_race)` to group the data by race/ethnicity
which is then passed on to the `summarise()` function to compute the
counts. Within the `summarise()` function, `count` is the name we want
to give to the column of the summary table and `n()` tells the program
to count the number of observations. In other words, we're telling the
program to count the number of obs. per group and put it into a column
named "count".

We can also group by two variables (e.g. race/ethnicity and age)

```{r}
df.nhis.2 %>% 
  group_by(r_race, r_age) %>% 
  summarise(count = n())

# EXERCISE: Using group_by() and summarise() compute the number of people with diabetes in each race group
df.nhis.2 %>%
  group_by(r_race, r_diab) %>%
  summarise(count = n())
```

Now that we have the counts, we naturally want to compute the frequency.
This is a bit trickier to do but straight forward once you get used to
it. Going back to the race/ethnicity and age example, we need to first
use `ungroup()` to reset the grouping. We than call `group_by()` again
with only the `r_race` variable such that we only group by
race/ethnicity. Now, when we use statistical summary functions such as
`n()`, `sum()`, `mean()`, `median()`, they will calculate the value for
each race/ethnicity group. To create a new column, recall that we use
`mutate()`. We can create a new column named `prop` by dividing the raw
count by the total count within each race/ethnicity group, computed by
`sum(count)` to obtain the frequency.

```{r}
df.nhis.2 %>%
  group_by(r_race, r_age) %>% 
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(r_race) %>%
  mutate(prop = count/sum(count) * 100)
```

## Introduction to Data Visualization with ggplot2

For your in-class projects this week you will be exploring the data sets
in order to gain a better understanding of them and to start developing
your research question. You will be presenting these preliminary results
to the implementation fellows, TAs, as well as Drs. Srinivasan, and
Palaniappan. To maximally capitalize our time together, you should make
sure that your data is presented with graphics that are easy to read and
understand. This section will cover the basics of\`ggplot2\` and how to
create some basic plots that you will be using heavily this week.

`ggplot2` is a R package for producing statistical graphics and is
written by [Hadley Wickham](http://hadley.nz), a giant for modern
statistical computing in R. Unlike other graphics packages, `ggplot2`
has an underlying
[grammar](https://www.springer.com/gp/book/9780387245447) which makes it
easy to learn and intuitive to work with. I will only be going over the
very basics here, but if you want to master \`ggplot2\` Dr. Wickham has
released his [book](https://ggplot2-book.org/introduction.html) for free
online.

### First Steps

![The 7 layers of
ggplot2](https://raw.githubusercontent.com/ShozenD/care-tutorials-2021/main/figures/ggplot-layered.png){width="367"}

When you create plots using \`ggplot2\` you want to think in terms of
"layers". The figure above gives you a very good image of each layer and
their ordering. Let's start with the three most important layers at the
bottom.

Every \`ggplot2\` plot has three key components:

1.  **data**. The foundation of your plot and always the first argument
    of the `ggplot` function.
2.  A set of **aesthetic mappings** between variables in the data and
    visual properties (e.g. x-axis, y-axis, color, shape, size).
    Aesthetic mappings are defined using the `aes` function.
3.  At least one geometric layer which describes how to render (i.e.
    draw) each observation. Layers are usually created with a `geom`
    function.

```{r}
head(iris)

ggplot(iris, aes(x = Sepal.Length, y = Petal.Length)) + 
  geom_point()

# EXERCISE: Create a scatter plot of Sepal.Length and Sepal.Width and color it by Species
```

Pay attention to the structure of program (i.e. function call): data and
aesthetic mappings are supplied to `ggplot()`, then geometric layers are
added with `+`. This is the basic pattern for every `ggplot2` program.

```{r}
# EXERCISE: Lets create some plots using df.nhis.2

# 1. A histogram of age. hint: Use geom_histogram()
ggplot(df.nhis.2, aes(x = age)) + 
  geom_histogram()

# 2. Boxplots with recoded race (r_race) on the x-axis and age on the y-axis. 
# hint: Use geom_boxplot()
ggplot(df.nhis.2, aes(x = r_race, y = age)) + 
  geom_boxplot()

# 3. A barplot for race counts (r_race). hint: geom_bar()
ggplot(df.nhis.2, aes(r_race)) + 
  geom_bar()
```

Now that we have the basics down, we can begin to use `ggplot` to
visualize our data. In a previous section you learned about how to make
bi-variate summary tables. For example, if we want to find the number of
respondents with diabetes within each race/ethnicity group we can easily
create a table by:

```{r}
table(df.nhis.2$r_race, df.nhis.2$r_diab)
```

However, when you present this data you want use plots not tables. To do
this we need to go through the following two steps.

1.  Count the number of obs. in each group.
2.  Create a stacked bar plot

Recall that we can count obs. in groups using the `group_by()` and
`summarise()` functions.

```{r}
df.diab <- df.nhis.2 %>% 
  group_by(r_race, r_diab) %>% 
  summarise(n = n())
```

We can then pass this data to ggplot2 for plotting. We want
race/ethnicity to be on the x-axis, the number of counts on the y-axis,
and we want to color the bars by diabetes status. Recall that to create
a bar plot we use the `geom_bar()` function. In order to create a
stacked bar plot, within the `geom_bar()` function we need to set
`position = "fill"` and `stat = "identity"` as such:

```{r}
ggplot(df.diab, aes(x = r_race, y = n, fill = r_diab)) + 
  geom_bar(
    position = "fill", 
    stat = "identity"
  )
```

Now, instead of absolute counts we can see the proportion of people with
diabetes in each race/ethnicity group! But we can't show this plot to
anyone yet because it doesn't have a title, the y-axis has the wrong
label, and the legend has a terrible name. We can customize the labels
using the `labs()` function.

```{r}
ggplot(df.diab, aes(x = r_race, y = n, fill = r_diab)) + 
  geom_bar(
    position = "fill", 
    stat = "identity"
  ) + 
  labs(
    x = "Race/Ethnicity",
    y = "Proportion",
    fill = "Diabetes Status"
  )
```

If you still feel like the plot is not to your liking you can make it
fancier by adding other options to `geom_bar()`, and by adding other
layers to the plot. We don't have time to cover everything today but
here is a example:

```{r}
ggplot(df.diab, aes(x = r_race, y = n, fill = r_diab)) + 
  geom_bar(
    width = 0.5,
    position = "fill", 
    stat = "identity"
  ) + 
  labs(
    x = "Race/Ethnicity",
    y = "Proportion",
    fill = "Diabetic"
  ) +
  scale_fill_manual(values = c("#488f31", "#de425b")) + 
  theme_minimal() + 
  theme(aspect.ratio = 1)
```
