---
title: "Introduction to Data Analysis in R (2)"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
library(tidyverse)
detach(package:plyr)
```

```{r message=FALSE}
df.nhis <- read_csv("./data/nhis_clean.csv")
```

In the last session we covered the basics of data wrangling
(pre-processing) with
`dplyr`and some ways of visualizing uni-variate and bi-variate data using `ggplot2`. This session will be about how you can discover and test **associations** (not to be confused with causation) in your data sets, which is an important intermediate step before progressing to statistical inference using regression models.

## Bivariate Data Visualization
### Scatter Plots
The first choice for visualizing bi-variate data is a scatter plot. However, in the case of NHIS most variables are categorical (nominal, ordinal, binary) and even if they are numerical, they are on a integer scale, which makes them unsuitable for scatter plots. As an example lets create a scatter plot with psychological distress (Kessler-6 scale) scores and neighborhood social cohesion scores.
```{r}
ggplot(df.nhis, aes(k6, nsc)) + 
  geom_point()

# We can mitigate the point overlap issue using geom_jitter(), which disperses the points randomly. However, as there are so many data points its still hard to gain any insights. 
ggplot(df.nhis, aes(k6, nsc)) + 
  geom_jitter(alpha = 0.1)
```

### Boxplots
A better way to visualize the relationship between psychological distress and neighborhood social cohesion would be to use box plots. Before we plug the data into `ggplot2`, we need to factorize the data (CSV files unfortunately cannot encode factor information so we need to do this every time we re-load data).
```{r}
df.nhis.md_nsc <- df.nhis %>% 
  filter(race %in% c("NHW", "Chinese", "Asian Indian", "Filipino")) %>%
  mutate(
    md = factor(md, levels = c("None", "Mild", "Moderate", "Severe")),
    race = factor(race, levels = c("NHW", "Asian Indian", "Chinese", "Filipino"))
  ) %>% 
  drop_na(nsc, md)

ggplot(df.nhis.md_nsc, aes(md, nsc)) + 
  geom_boxplot()

ggplot(df.nhis.md_nsc, aes(md, nsc, fill = race)) + 
  geom_boxplot(width = 0.3) + 
  # facet_wrap(~race, nrow = 1, ncol = 4) + 
  theme()
```
So, it seems like mental distress has a negative correlation with neighborhood social cohesion, especially for non-Hispanic Whites and Asian Indians (in comparison to Chineses and Filipinos).

### Barplots
```{r}
nsc.tertiles <- quantile(df.nhis.md_nsc$nsc, probs = c(.33, .66))

df.nhis.md_nsc <- df.nhis.md_nsc %>% 
  mutate(
    rnsc = case_when(
      nsc < nsc.tertiles[1] ~ "low",
      nsc < nsc.tertiles[2] ~ "medium",
      nsc >= nsc.tertiles[2] ~ "high"
    )
  ) %>% 
  mutate(rnsc = factor(rnsc, levels = c("low", "medium", "high"))) # You can't put factor in the same mutate

df.md_nsc.sum <- df.nhis.md_nsc %>% 
  group_by(rnsc, md) %>% 
  dplyr::summarise(
    N = n()
  ) %>% 
  ungroup(md) %>% 
  mutate(
    prop = N/sum(N)
  )

ggplot(df.md_nsc.sum, aes(x = rnsc, y = prop, fill = md)) + 
  geom_bar(
    position = "fill", 
    stat = "identity"
  )
```

## Bivariate Association Tests
### Correlation
```{r}
cor.test(~ k6 + nsc, df.nhis.md_nsc)

cor.test(~ k6 + nsc, data = df.nhis.md_nsc, method = "spearman")

cor.test(~k6 + nsc, filter(df.nhis.md_nsc, race == "NHW"))
```

### t-test
```{r}
```


### Chi-squared test
```{r}
table(df.nhis.md_nsc$md, df.nhis.md_nsc$rnsc) %>% chisq.test()
```

### ANOVA


## Directed Acyclic Graphs (DAGs)
### Superious Correlations
https://www.tylervigen.com/spurious-correlations

Bi-variate data visualization (NHIS)

Bi-variate tests for association

How to check for con-founders in R (put in regression) process of
identifying con-founders How to add an interaction term

Assumptions for linear and logit regression Outcome variables cannot be
too skewed

qq-plot

equal residual variance

```{r, results = F}
NHISdataAdult2018$DEP_1 # How often do you feel depressed
NHISdataAdult2018$DEP_2 # do you take dep medicine
NHISdataAdult2018$DEP_3R #how depressed were you the last time you felt depressed
NHISdataAdult2018$ALDURA17 #Duration of depresion
NHISdataAdult2018$AFLHCA17 #depression causes difficulty with activity
NHISdataAdult2018$ASISAD # so sad nothign cheers you up
NHISdataAdult2018$ASIHOPLS #hopeless
NHISdataAdult2018$ASIWTHLS #worthless
NHISdataAdult2018$ASINERV #felt nervous
NHISdataAdult2018$ASIRSTLS #felt restless
NHISdataAdult2018$ASIEFFRT #felt everything was an effort
NHISdataAdult2018$ASIMUCH #feelings interfered with life

NHISdataAdult2018$AHCAFYR2 # couldnt afford mental health counseling
NHISdataAdult2018$AHCSYR1 # have you seen mental health practicioner in past 12 months
```

Tests!!!

Correlation: Find association between variables based on r value
(closest to -1 or 1 indicated stronger association, and close to 0 means
little association). Can find the association between both catgeorical
and continuous variables. -pearson - parametric, linear
relationships(depends on distribution - x and y should have normal
distribution) -Use for two continuous variables -kendall and spearman -
non-parametric -can work with continuous or categorical variables

    Cor(), cor.test(), corrplot.mixed()

T test - contiuous -standard 2 sample t test: use to determine whether
difference in means between two groups -pairwise t test: multiple t
tests between all possible combinations of groups

Chi Squared - categorical -Chi square test of independence: compares two
variables in a contingency table to see if they are related. In a more
general sense, it tests to see whether distributions of categorical
variables differ from each another -small chi square result means
obsevred data fits expected value well - relationship!

ANOVA - analysis of variance between multiple groups -Like the 2 sample
t test, but can be for multiple groups! Check to see if mean differers
between all the groups

```{r}
# Correlations with pearson - continuous
#onlyAsianAndWhite$ALDURA17

hist(onlyAsianAndWhite$ALDURA17)

# Clean the histogram
duraCleaned <- onlyAsianAndWhite[!(is.na(onlyAsianAndWhite$ALDURA17)), ]
duraCleaned <- onlyAsianAndWhite[onlyAsianAndWhite$ALDURA17 < 97, ]
summary(duraCleaned$ALDURA17)
hist(duraCleaned$ALDURA17, col = "blue")
hist(log(duraCleaned$ALDURA17+1), col = "blue")
boxplot(duraCleaned$ALDURA17 ~ duraCleaned$raceRecoded, col = c("blue", "red", "pink", "green"))

#Basic visualization as a histogram
hist(duraCleaned$ASISLEEP)

#Cleaned up histogram
sleepCleaned <- duraCleaned[duraCleaned$ASISLEEP < 20, ]
hist(sleepCleaned$ASISLEEP, col = "blue")

plot(sleepCleaned$ALDURA17, sleepCleaned$ASISLEEP, pch = 19, col = "blue")

lm1 <-  lm(sleepCleaned$ASISLEEP ~ sleepCleaned$ALDURA17)
plot(sleepCleaned$ALDURA17, sleepCleaned$ASISLEEP, pch = 19, col = "blue")
abline(lm1$coef, col = "red", lwd = 3)

#cor(log(sleepCleaned$ALDURA17+1), sleepCleaned$ASISLEEP, use = "na.or.complete", method = "pearson")
cor(sleepCleaned$ALDURA17, sleepCleaned$ASISLEEP, use = "na.or.complete", method = "pearson")

cor.test(sleepCleaned$ALDURA17, sleepCleaned$ASISLEEP, method = "pearson")
```

```{r}
## Correlations with spearman - continous v continuous or continous v categorical

# continuous vs categorical 
asirstls <- duraCleaned[!(duraCleaned$ASIRSTLS == 7 | duraCleaned$ASIRSTLS == 8 | duraCleaned$ASIRSTLS == 9), ]

plot(asirstls$ALDURA17 ~ asirstls$ASIRSTLS, pch = 19, col = "blue")

plot(jitter(asirstls$ALDURA17) ~ jitter(asirstls$ASIRSTLS), pch = 19, col = "blue")

lm1 <-  lm(asirstls$ASIRSTLS ~ asirstls$ALDURA17)
plot(jitter(asirstls$ALDURA17) ~ jitter(asirstls$ASIRSTLS), pch = 19, col = "blue")
abline(lm1$coef, col = "red", lwd = 3)

(cor <- cor(asirstls$ALDURA17, asirstls$ASIRSTLS, use = "na.or.complete", method = "spearman"))

cor.test(asirstls$ALDURA17, asirstls$ASIRSTLS, method = "spearman", exact = F)

#categorical vs categorical
dep_1 <- asirstls[!(asirstls$DEP_1 == 7 | asirstls$DEP_1 == 8 | asirstls$DEP_1 == 9), ]

plot(jitter(dep_1$DEP_1) ~ jitter(dep_1$ASIRSTLS), pch = 19, col = "blue")

lm1 <-  lm(dep_1$ASIRSTLS ~ dep_1$DEP_1)
plot(jitter(dep_1$DEP_1) ~ jitter(dep_1$ASIRSTLS), pch = 19, col = "blue")
abline(lm1$coef, col = "red", lwd = 3)

(cor <- cor(dep_1$DEP_1, dep_1$ASIRSTLS, use = "na.or.complete", method = "spearman"))
cor.test(dep_1$DEP_1, dep_1$ASIRSTLS, method = "spearman", exact = F)
```

```{r fig.height = 5, fig.width = 5}
#CorPlots

cp <- onlyAsianAndWhite[,c("DEP_1","DEP_2","DEP_3R","ALDURA17","ASISAD","ASIHOPLS","ASINERV","ASIRSTLS","ASIEFFRT","ASIMUCH","AHCAFYR2","AHCSYR1")]
cp <- cp[complete.cases(cp), ]

#Basic plot
round(cor(cp), 2)
corrplot(cor(cp), method = "ellipse")

#First, get a matrix of correlation tests.
sigcorr <- cor.mtest(cp, conf.level = .95) #test of of correlations statistically significant in matrix
#sigcorr

#Make cool plot
corrplot.mixed(cor(cp), lower.col = "black", upper = "ellipse", tl.col = "black", number.cex = .7,order = "hclust", tl.pos = "lt", tl.cex=.7, p.mat = sigcorr$p, sig.level = .05) 

#Male even cooler plot
chart.Correlation(cp, histogram = TRUE, pch = 19)
```

```{r}
# what to watch out for!

cor(cp$DEP_2, cp$DEP_3R, use = "na.or.complete")
cor.test(cp$DEP_2, cp$DEP_3R, method = "spearman", exact = F)

plot(cp$DEP_2, cp$DEP_3R)

c1 <- cp[!(cp$DEP_2 == 7 | cp$DEP_2 == 8 | cp$DEP_2 == 9), ]
c2 <- c1[!(c1$DEP_3R == 7 | c1$DEP_3R == 8 | c1$DEP_3R == 9), ]

cor(c2$DEP_2, c2$DEP_3R, use = "na.or.complete")
cor.test(c2$DEP_2, c2$DEP_3R, method = "spearman", exact = F)

```

T Test

```{r}
# 2 Sample T test
duraCleaned$sexRecoded <- Recode(duraCleaned$SEX, '"1" = "Male"; "2" = "Female"')

boxplot(duraCleaned$ALDURA17 ~ duraCleaned$sexRecoded, col = c("pink", "blue"))
t.test(duraCleaned$ALDURA17 ~ duraCleaned$sexRecoded) 
```

```{r}
#pairwise t test 

pairwise.t.test(duraCleaned$ALDURA17, duraCleaned$raceRecoded, p.adjust = "bonferroni")
```

```{r}
#ANOVA

sds <- tapply(duraCleaned$ALDURA17, duraCleaned$raceRecoded, sd)
round(max(sds)/min(sds), 1)

boxplot(as.numeric(duraCleaned$ALDURA17) ~ as.factor(duraCleaned$raceRecoded), col = "yellow", xlab = "Race")
means <- tapply(duraCleaned$ALDURA17, duraCleaned$raceRecoded, mean) 
points(means, col = "red", pch = 19)
text(x = c(1:4), y = means + 5, labels = round(means,2))

mod1 <-  aov(as.numeric(duraCleaned$ALDURA17) ~ as.factor(duraCleaned$raceRecoded))
summary(mod1)

TukeyHSD(mod1)

par(mar = c(5,13,2,2))
plot(TukeyHSD(mod1), las = 1)

qqPlot(duraCleaned$ALDURA17, pch = 19)

qqPlot(log(duraCleaned$ALDURA17+1), pch = 19)
```

```{r}
#Chi squared
  #null is that the two are independent

t = table(dep_1$DEP_1, dep_1$ASIRSTLS)

chisq.test(t)
```
